%%% The main file. It contains definitions of basic parameters and includes all other parts.

%% Settings for single-side (simplex) printing
% Margins: left 40mm, right 25mm, top and bottom 25mm
% (but beware, LaTeX adds 1in implicitly)
\documentclass[12pt,a4paper]{report}
\setlength\textwidth{145mm}
\setlength\textheight{247mm}
\setlength\oddsidemargin{15mm}
\setlength\evensidemargin{15mm}
\setlength\topmargin{0mm}
\setlength\headsep{0mm}
\setlength\headheight{0mm}
% \openright makes the following text appear on a right-hand page
\let\openright=\clearpage

%% Settings for two-sided (duplex) printing
% \documentclass[12pt,a4paper,twoside,openright]{report}
% \setlength\textwidth{145mm}
% \setlength\textheight{247mm}
% \setlength\oddsidemargin{14.2mm}
% \setlength\evensidemargin{0mm}
% \setlength\topmargin{0mm}
% \setlength\headsep{0mm}
% \setlength\headheight{0mm}
% \let\openright=\cleardoublepage

%% Generate PDF/A-2u
\usepackage[a-2u]{pdfx}

%% Character encoding: usually latin2, cp1250 or utf8:
\usepackage[utf8]{inputenc}

%% Prefer Latin Modern fonts
\usepackage{lmodern}

%% Further useful packages (included in most LaTeX distributions)
\usepackage{amsmath}        % extensions for typesetting of math
\usepackage{amsfonts}       % math fonts
\usepackage{amsthm}         % theorems, definitions, etc.
\usepackage{bbding}         % various symbols (squares, asterisks, scissors, ...)
\usepackage{bm}             % boldface symbols (\bm)
\usepackage{graphicx}       % embedding of pictures
\usepackage{fancyvrb}       % improved verbatim environment
\usepackage{natbib}         % citation style AUTHOR (YEAR), or AUTHOR [NUMBER]
\usepackage[nottoc]{tocbibind} % makes sure that bibliography and the lists
			    % of figures/tables are included in the table
			    % of contents
\usepackage{dcolumn}        % improved alignment of table columns
\usepackage{booktabs}       % improved horizontal lines in tables
\usepackage{paralist}       % improved enumerate and itemize
\usepackage{xcolor}         % typesetting in color

\usepackage{algpseudocode}
\usepackage{algorithm}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}



%%% Basic information on the thesis

% Thesis title in English (exactly as in the formal assignment)
\def\ThesisTitle{Rotation-equivariant convolutional neural network for design of visual prosthetic stimulation protocol}

% Author of the thesis
\def\ThesisAuthor{Martin Picek}

% Year when the thesis is submitted
\def\YearSubmitted{2022}

% Name of the department or institute, where the work was officially assigned
% (according to the Organizational Structure of MFF UK in English,
% or a full name of a department outside MFF)
\def\Department{Department of Software and Computer Science Education}

% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Department of Software and Computer Science Education}

% Thesis supervisor: name, surname and titles
\def\Supervisor{Mgr. Ján Antolík, Ph.D.}

% Supervisor's department (again according to Organizational structure of MFF)
\def\SupervisorsDepartment{Department of Software and Computer Science Education}

% Study programme and specialization
\def\StudyProgramme{Computer Science}
\def\StudyBranch{Artificial Intelligence}

% An optional dedication: you can thank whomever you wish (your supervisor,
% consultant, a person who lent the software, etc.)
\def\Dedication{%
First and foremost, I want to thank my supervisors. With excellent teaching skills and a friendly attitude, Mgr. Ján Antolík, Ph.D. inspired me and many other students to work in the field of computational neuroscience. I was fortunate to have such a patient supervisor willing to devote a great deal of time to his students. I am grateful for all the invaluable knowledge that I was given from him.

Luca Baroni was the person to whom I could come for any help. Despite his very tight schedule, he offered me every spare minute of his time to support me in writing this thesis. I am glad that during this journey, we became not only good colleagues but, most importantly, very good friends.

Furthermore, I want to acknowledge Dávid, David, Tomáš, and Radovan, who supported me throughout writing this thesis. Thank you, Vendulka, MzM, and other friends, for giving me all the emotional care I needed.

And most importantly, I want to thank my family, for the never-ending love they give me.

}

% Abstract (recommended length around 80-200 words; this is not a copy of your thesis assignment!)
\def\Abstract{%
Neighboring neurons in the primary visual cortex (V1), the first cortical area processing visual information, are selective to stimuli presented in neighboring positions of the visual field with a specific edge orientation. In this way, they form the so-called retinotopic and orientation maps of V1. Due to the absence of high-resolution cortical stimulation devices, vision restoration through prosthetic implants in V1 has not yet taken advantage of the orientation maps. However, the availability of cortical implants with stimulation resolution high enough to target separate orientation columns can be anticipated soon. 

Since other stimulus features are also encoded in the cortex, such as color, size, or phase, but cannot be reliably engaged even by high-resolution stimulation, in this thesis, we ask the question of how well can visual stimuli be encoded in V1 if only orientation and position preference is known. To address this question, we propose a deep neural network (DNN) providing a scalar neural activity descriptor for any targeted cortical location and multiple different orientations. This is achieved by employing a rotation-equivariant convolutional neural network (reCNN) with the last layer having only one channel for each orientation, returning the desired three-dimensional feature tensor. A specialized readout estimates the neurons’ positions and their orientation preference, using them to yield one value per neuron from the core’s last layer, from which the neural response is predicted. These scalar features might serve as an input for the future stimulation protocol.

The proposed network was trained both on an experimental and a synthetic dataset. The synthetic dataset generated by Antolík et al. model of a cat’s V1 provided locations of neurons along with their orientation preference. This was used to examine the DNN model’s precision in both neural response prediction and neural location and preferred orientation estimates. 

This work proposes a DNN model yielding a 0.8327 correlation fraction with respect to the control model. Moreover, with the proposed reCNN model achieving sufficient performance, we were able to reconstruct the in-silico orientation maps. This promising result suggests that targeting cortical columns of neurons matching in orientation and position preference might improve information encoding via prosthetic stimulation.
}

% 3 to 5 keywords (recommended), each enclosed in curly braces
\def\Keywords{%
{deep neural networks}, {computational neuroscience}, {rotation-equivariant CNN}, {convolutional neural network}, {cortical prosthetics}, {stimulation protocol}
}

%% The hyperref package for clickable links in PDF and also for storing
%% metadata to PDF (including the table of contents).
%% Most settings are pre-set by the pdfx package.
\hypersetup{unicode}
\hypersetup{breaklinks=true}

% Definitions of macros (see description inside)
\include{macros}

% Title page and various mandatory informational pages
\begin{document}
\include{title}

%%% A page with automatically generated table of contents of the bachelor thesis

\tableofcontents

%%% Each chapter is kept in a separate file
\include{introduction}
\include{visual_system}
\include{deep_learning}
\include{related_work}
\include{methodology}
\include{experiments}
\include{discussion}
\include{conclusion}

%%% Bibliography
\include{bibliography}

%%% Figures used in the thesis (consider if this is needed)
\listoffigures

%%% Tables used in the thesis (consider if this is needed)
%%% In mathematical theses, it could be better to move the list of tables to the beginning of the thesis.
\listoftables

%%% Abbreviations used in the thesis, if any, including their explanation
%%% In mathematical theses, it could be better to move the list of abbreviations to the beginning of the thesis.
\chapwithtoc{List of Abbreviations}

\begin{description}
	
	\item[DNN] Deep neural network
	\item[CNN] Convolutional neural network
	\item[reCNN] Rotation-equivariant convolutional neural network
	\item[LGN] Lateral geniculate nucleus
	\item[V1] Primary visual cortex.
	\item[val] Validation (dataset)	
	\item[test] Test (dataset)
	\item[corr] Correlation	
	\item[wrt.] with respect to
	\item[MLE] Most likelihood estimation.

\end{description}

%%% Attachments to the bachelor thesis, if any. Each attachment must be
%%% referred to at least once from the text of the thesis. Attachments
%%% are numbered.
%%%
%%% The printed version should preferably contain attachments, which can be
%%% read (additional tables and charts, supplementary text, examples of
%%% program output, etc.). The electronic version is more suited for attachments
%%% which will likely be used in an electronic form rather than read (program
%%% source code, data files, interactive charts, etc.). Electronic attachments
%%% should be uploaded to SIS and optionally also included in the thesis on a~CD/DVD.
%%% Allowed file formats are specified in provision of the rector no. 72/2017.
\appendix
\chapter{Attachments}

\section{reCNN-visual-prosthesis}\label{github_main}
A repository with all the source code of our solution. The repository can also be found at \url{https://github.com/mpicek/reCNN_visual_prosthesis"}.


\section{csng-dl-docker-image}\label{docker_image}
A repository with a Docker image for running the training environment. Also available online at \url{https://github.com/mpicek/csng_dl_docker_image}.



\openright
\end{document}
